Nail: A practical interface generator for data formats

### 2014-11-25 paper/login/design.tex
More typowrangling

We will use DNS packets as a leading example, as defined in RFC 1035.  Each DNS packet consists of a
header, a set of question records, and a set of answer records. Domain
names in both queries and answers are encoded as a sequence of labels,
terminated by a zero byte.  Labels are Pascal-style strings, consisting
of a length field followed by that many bytes comprising the label.
\label{s:design}
\subsection{Basic Data formats}

\input{fig-dnsfull}
Let's step through a simplified Nail grammar DNS
packets, shown in  Figure~\ref{fig:dns-full}.  For this grammar, Nail produces the type declarations
shown in Figure~\ref{fig:dns-full-struct}, and the parser and
generator functions shown in Figure~\ref{fig:dns-full-api}.
Nail grammars are reuseable between applications and we will use this grammar to implement
both a DNS server and client, which previously would have had two separate hand-written parsers,
leading to bugs such as  the Android Master Key.

>>>>

As a motivating example, we will parse DNS packets, as defined in RFC 1035. Each DNS packet consists of a
header, a set of question records, and a set of answer records. Domain
names in both queries and answers are encoded as a sequence of labels,
terminated by a zero byte.  Labels are Pascal-style strings, consisting
of a length field followed by that many bytes comprising the label.
\label{s:design}
\subsection{Basic Data formats}

\input{fig-dnsfull}
Let's step through a simplified Nail grammar for DNS
packets, shown in  Figure~\ref{fig:dns-full}.  For this grammar, Nail produces the type declarations
shown in Figure~\ref{fig:dns-full-struct}, and the parser and
generator functions shown in Figure~\ref{fig:dns-full-api}.
Nail grammars are reuseable between applications and we will use this grammar to implement
both a DNS server and client, which previously would have had two separate hand-written parsers,
leading to bugs such as  the Android Master Key.

### 2014-11-25 paper/login/intro.tex
More typowrangling

Furthermore, some parser generators are cumbersome to use when parsing binary data for several
reasons. First, many parser generators don't directly data structures, but call semantic actions
that the you write to construct a data structure. Therefore,you must
describe the format up to three times -- in the grammar, the data structure and the semantic
actions. Second, most parser generators only address parsing inputs, so you'd have to manually
construct outputs. Some parser generators, such as 
Boost.Spirit  allow generating output, but require you to write another set of semantic actions.

We address these challenges with Nail, a new parser generator for binary formats.
First, Nail grammars describe not only a format, but also a data type to represent it within the program.
Therefore you don't have to write semantic actions and type declarations and you can no longer syntactic validation and semantic processing.
Second, Nail will also take care of generating output from the same data type, without requiring you
to write more risky code or think about the format again.

Third,  Nail introduces two abstractions, \emph{dependent fields} and
\emph{transformations}, to elegantly handle problematic structures,
such as offset fields or checksums.  Dependent fields capture fields
in a protocol whose value depends in some way on the value or layout
of other parts of the format; for example, offset or length fields,
which specify the position or length of another data structure, fall
in this category.  Transformations allow you to escape the
generated code to modify the raw data and interact with dependent fields
in a controlled manner.

>>>>

Furthermore, some parser generators are cumbersome to use when parsing binary data for several
reasons. First, many parser generators don't produce convenient data structures, but call
\emph{semantic actions} that you have to write to build up a data structure your program can use. Therefore,you must
describe the format up to three times -- in the grammar, the data structure and the semantic
actions. Second, most parser generators only address parsing inputs, so you have to manually
construct outputs. Some parser generators, such as 
Boost.Spirit  allow generating output, but require you to write another set of semantic actions.

We address these challenges with Nail, a new parser generator for binary formats.
First, Nail grammars describe not only a format, but also a data type to represent it within the program.
Therefore you don't have to write semantic actions and type declarations and you can no longer syntactic validation and semantic processing.
Second, Nail will also generate output from this data type, without requiring you
to write more risky code or giving you a chance to introduce ambiguity.

Third,  Nail introduces two abstractions, \emph{dependent fields} and
\emph{transformations}, to elegantly handle problematic structures,
such as offset fields or checksums.  Dependent fields capture fields
in a protocol whose value depends in some way on the value or layout
of other parts of the format; for example, offset or length fields,
which specify the position or length of another data structure, fall
in this category.  Transformations allow you to write plug-ins, allowing your programs to handle complicated structures, while
keeping Nail itself small, yet flexible.
